---
title: "Untitled"
author: "Julia Christensen"
date: "February 13, 2019"
output: html_document
---

***************************************************
# 1) SET UP 

## 1.1. General Set Up 
```{r}
### Clear global environment
rm(list=ls()) 

### Download Libraries 
library(pacman)
p_load(tidyverse, data.table, rmarkdown, knitr, tinytex, magrittr,
       Matrix,
       haven,
       sjmisc, #for adding dummies (https://stackoverflow.com/questions/49276914/mutating-dummy-variables-in-dplyr)
       Matching, permute, car, sandwich, 
       lfe, AER, ivmodel, # Added for 2SLS
       estimatr, purrr,
       rlist,
       stargazer, xtable,
       viridis
       )

# install.packages("haven")
# library(haven)

### Set file paths
# source("../Set_Up/file_paths.R")

### Download visualisation formatting
source("../Set_Up/visualisation.R")

### Download functions
source("../Functions/fun_peng_iv.R")
source("../Functions/fun_iv_analysis.R")

```


## 1.2 Load Data
```{r, eval=FALSE}
# Clear global environment (NOT functions & ggplot theme)
rm(list=setdiff(ls(), c(lsf.str(), "theme_ed"))) 

# Select data
agg_lvl  <- "cz"
dis_type <- "f"
fema_sub <- FALSE

if(dis_type=="f") dis_name <- "flood"
if(dis_type=="t") dis_name <- "tornado"

# Download data
if(fema_sub==TRUE){
  dat <- load_clean_fun(agg_lvl = agg_lvl, dis_type = dis_type, 
                        fp=paste0("../Data/storm_",dis_type,"_1to1_",agg_lvl,"_merged.RData"))
} else{
  dat <- load_clean_fun(agg_lvl = agg_lvl, dis_type = dis_type, 
                        fp=paste0("../Data/storm_1to1_",agg_lvl,"_merged.RData"))
}

# Make Z Binary 
# dat%<>% mutate_at(vars(contains("np_event")), funs(ifelse(.>0,1,0)))

# Add totals if cz
if(agg_lvl=="cz"){
  dat %<>% mutate(deaths_total   = DEATHS_DIRECT_num + DEATHS_INDIRECT_num,
                  injuries_total = INJURIES_DIRECT_num + INJURIES_INDIRECT_num)
}
```

***************************************************
# (2) ANALYSIS

Comparisons to run:
- All FEMA disasters (match_fema) vs. flood/tornado disasers (match_fema_f/t)
- NYT hits with state vs. cz search criteria 
- Binary vs. continuous z
- np vs np_event 
- np_event subtypes
- Length of lead period 
- No X vs. FE only vs. X 
- FE
  * no FE
  * state vs. cz
  * m, y vs m-y
  * m-y-geo

Analysis:
- Lin's estimator 
- Confirm Lin's estimator = 2SLS results 
- Fieller-Anderson-Rubin confidence interval
- Fisher randomization 


```{r, eval=FALSE}
hist(dat$days_of_disaster)
hist(dat$deaths_total)
hist(dat$injuries_total)

hist(dat$nyt_hit_flag_st_b4decl_sub)
hist(dat$nyt_hits_total_st_b4decl_sub)

hist(dat$np_event_sum_b4decl_sub)

hist(dat$np_event_lead60)

hist(dat$match_fema_sub)

dat %<>% mutate(days_b4_decl = as.numeric(date_declaration_sub - new_date_beg))

dat %>% dplyr::select(days_b4_decl) %>% head(50) %>% t()

dat$np_event_sum_b4decl %>% is.na() %>% sum()

dat %>% 
  filter(days_b4_decl > -10 & days_b4_decl < 100) %>% 
  dplyr::select(days_b4_decl) %>%
  mutate(days_b4_decl = as.numeric(days_b4_decl) ) %>% 
  .$days_b4_decl %>% hist()

hist(dat$days_b4_decl)


dat %>%
  mutate(flag1 = ifelse(np_event_sum_b4decl >0, 1, 0),
         flag2 = ifelse(nyt_hits_total_st_b4decl >0, 1, 0),
         flag3 = ifelse(flag1==1 & flag2==1, 1, 0)) %>%
  summarise(pct_np = mean(flag1),
            pct_nyt = mean(flag2),
            pct_np_nyt = mean(flag3))

hist(dat$np_event_lead60)

```


## 2.1. Functions 

### 2.1.0. Load Data 
```{r}
load_dat_for_analysis <- function(agg_lvl, dis_type = "f"){
  # Set options
    if(dis_type=="f") dis_name <- "flood"
    if(dis_type=="t") dis_name <- "tornado"
  # Download data
    dat <- load_clean_fun(agg_lvl = agg_lvl, dis_type = dis_type) 
  # Add totals if cz
    if(agg_lvl=="cz"){
      dat %<>% mutate(deaths_total   = DEATHS_DIRECT_num + DEATHS_INDIRECT_num,
                      injuries_total = INJURIES_DIRECT_num + INJURIES_INDIRECT_num)
    }
  return(dat)
}
```

### 2.1.1. Load Data (1to1)
```{r}
load_dat_for_analysis_1to1 <- function(fema_sub, agg_lvl, dis_type = "f"){
  # Set options
    if(dis_type=="f") dis_name <- "flood"
    if(dis_type=="t") dis_name <- "tornado"
  # Download data
    if(fema_sub==TRUE){
      dat <- load_clean_fun(agg_lvl = agg_lvl, dis_type = dis_type, 
                            fp=paste0("../Data/storm_",dis_type,"_1to1_",agg_lvl,"_merged.RData"))
    } else{
      dat <- load_clean_fun(agg_lvl = agg_lvl, dis_type = dis_type, 
                            fp=paste0("../Data/storm_1to1_",agg_lvl,"_merged.RData"))
    }    
  # Add totals if cz
    if(agg_lvl=="cz"){
      dat %<>% mutate(deaths_total   = DEATHS_DIRECT_num + DEATHS_INDIRECT_num,
                      injuries_total = INJURIES_DIRECT_num + INJURIES_INDIRECT_num)
    }
  return(dat)
}
```

### 2.1.2. Gen lists for latex table
```{r}
### Gen SE for Table
# Single calc
  calc_single_se <- function(res)
  {
    # Save coefficient table & col names
      b_tbl <- res %>% summary(robust=TRUE) %>% coef()
      b_cols <- b_tbl %>% colnames()
    # Return SE
      if("Robust s.e" %in% b_cols) res_se <- b_tbl %>% .[,"Robust s.e"]
      if("Cluster s.e." %in% b_cols) res_se <- b_tbl %>% .[,"Cluster s.e."]
    return(res_se)  
  }
# Whole vector
  calc_se_vec <- function(res_list) map(res_list, calc_single_se)
```

```{r}
### Gen 1st Stage Coef (tauD)
# Single calc
  calc_single_td <- function(res, z, rnd=3)
  {
    res$stage1$coefficients[z,1] %>% round(rnd)
  }
# Whole vector
  calc_td_vec <- function(res_list, z, rnd=3)
  {
    map(res_list, calc_single_td, z=z, rnd=rnd) %>% unlist() %>% 
      unlist() %>% unname() %>%
      round(rnd)
  }
```

```{r}
### Gen tauY
# Single calc
  calc_single_ty <- function(res, z, d, rnd=3)
  {
    # Store final 2SLS beta for D
      d_txt <- paste0("`",d,"(fit)`")
      b <- res$coefficients[d_txt,]
    # Store 1st stage (tauD)  
      td <- calc_single_td(res, z)
    # Calc tauY
      b*td %>% round(rnd)
  }
# Whole vector
  calc_ty_vec <- function(res_list, z, d, rnd=3) 
  {
    map(res_list, calc_single_ty, z=z, d=d, rnd=rnd) %>% 
      unlist() %>% unname() %>%
      round(rnd)
  }
```

### 2.1.3. Weak IV Tests  
```{r}
### F-stat from 1st stage 
# Single calc
  get_s1_fstat_felm <- function(res, rnd=3, robust=TRUE)
  {
    # Get robust 
      if(robust==TRUE) f <- res$stage1$rob.iv1fstat[[1]]['F'] %>% round(rnd)
    # Get regular 
      if(robust==FALSE) f <- res$stage1$iv1fstat[[1]]['F'] %>% round(rnd)    
    # Return
      return(f)  
  }
# Whole vector
  get_s1_fstat_vec <- function(res_list, rnd=3, robust=TRUE) 
  {
    map(res_list, get_s1_fstat_felm, rnd=rnd, robust=robust) %>% 
      unlist() %>% unname() %>%
      round(rnd)
  }
```
Note: when there are multiple z values, a classic weak IV wald test does not equal the value from felm (see https://www.r-bloggers.com/detecting-weak-instruments-in-r/ for classic test). However, when there is only one z variable, the two methods seem to agree. Because lm() won't run for the FE models, I only tested this for the model with X and no FE. 
```{r, eval=FALSE}
### Testing 
# do regressions for partial F-tests
# first-stage:
lm_txt <- as.formula(paste0(d_txt," ~ 1 + ",X_txt,"+",z_txt)) %>% print()
fs = lm(lm_txt, data = dat)
# null first-stage (i.e. exclude IVs):
lm_txt_noIV <- as.formula(paste0(d_txt," ~ 1 + ",X_txt)) %>% print()
fn = lm(lm_txt_noIV, data = dat)

# simple F-test
waldtest(fs, fn)$F[2]
print("compare to felm")
get_s1_fstat_felm(lm2, robust=FALSE)

# F-test robust to heteroskedasticity
waldtest(fs, fn, vcov = vcovHC(fs, type="HC0"))$F[2]
print("compare to felm")
get_s1_fstat_felm(lm2, robust=TRUE)
lm2 %>% condfstat(type="robust")

sapply(lm2$stage1$lhs, function(lh) lfe::waldtest(lm2$stage1, ~np_olympics_lead20|np_sport_lead20|np_death_lead20, lhs=lh)) %>% t()

sapply(lm2$stage1$lhs, function(lh) lfe::waldtest(lm2$stage1, ~np_olympics_lead20|np_sport_lead20|np_death_lead20, lhs=lh, type="robust", r=0)) %>% t()

paste0("~",paste(p[["z"]], collapse="|"))
```

Compute conditional F statistic for weak instruments in an IV-estimation with multiple endogenous variables using lfe::condfstat(). Based on Sanderson, E. and F. Windmeijer (2014) A weak instrument F-test in linear IV models with multiple endogenous variables, Journal of Econometrics, 2015. http://www.sciencedirect.com/science/article/pii/S0304407615001736.
```{r}
### lfe::condfstat()
# Single calc
  get_s1_condfstat_felm <- function(res, rnd=3, robust=TRUE)
  {
    # Get robust 
      if(robust==TRUE) f <- res %>% condfstat(type="robust")
    # Get regular 
      if(robust==FALSE) f <- res %>% condfstat(type="iid")
    # Return
      return(f)  
  }
# Whole vector
  get_s1_condfstat_vec <- function(res_list, rnd=3, robust=TRUE) 
  {
    map(res_list, get_s1_condfstat_felm, rnd=rnd, robust=robust) %>% 
      unlist() %>% unname() %>%
      round(rnd)
  }
# Note: May want to add quantiles in future 
# condfstat(est, quantiles=c(0.05, 0.95))
```

### 2.1.4. Gen Reg Results
```{r}
gen_general_table_contents <- function(p)
{
  # IV Only (1)
    lm1 <- p[["dat"]] %>% 
      run_tsls_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = NULL,
                   geo_fe  = NULL,
                   time_fe = NULL )
  # Add FE (3)
    lm3 <- dat %>% 
      run_tsls_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = NULL,
                   geo_fe  = p[["geo_fe"]],
                   time_fe = p[["time_fe"]])
  # Add X (2)
    lm2 <- dat %>%
      run_tsls_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = p[["x_vars"]],
                   geo_fe  = NULL,
                   time_fe = NULL)  
  # Add FE + X (5)
    lm5 <- dat %>%
      run_tsls_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = p[["x_vars"]],
                   geo_fe  = p[["geo_fe"]],
                   time_fe = p[["time_fe"]])
  # Add FE + X + Clustering (7)
    lm7 <- dat %>%
      run_tsls_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = p[["x_vars"]],
                   geo_fe  = p[["geo_fe"]],
                   time_fe = p[["time_fe"]],
                   cluster_var = p[["cluster_var"]]) 
    
  return(list("lm1"=lm1, "lm2"=lm3, "lm3"=lm2, "lm4"=lm5, "lm5"=lm7))
}
```

### 2.1.5. Gen Table
```{r}
gen_general_table <- function(p, output_type="text")
{
  # Generate IV results  
    res_list <- gen_general_table_contents(p)
  # Track model choices
    extra_lines <- list(c("Covariates",  "", "", "X", "X", "X"),
                        c("Time FE",     "", "X", "", "X", "X"), 
                        c("Geo FE",      "", "X", "", "X", "X"),
                        c("Cluster SE",  "", "", "", "", "X")
                        )
  # Add tauD and tauY if only one var in model  
    if(length(p[["z"]])==1){
      extra_lines <- c(extra_lines,
                       list(c("tauD", calc_td_vec(res_list, p[["z"]]))),
                       list(c("tauY", calc_ty_vec(res_list, p[["z"]], p[["d"]]))),
                       # list(c("Stage1 F", get_s1_fstat_vec(res_list, robust=FALSE))),
                       list(c("Stage1 F (robust)", get_s1_fstat_vec(res_list, robust=TRUE))),
                       # list(c("Stage1 Conditional F", get_s1_condfstat_vec(res_list, robust=FALSE))),
                       list(c("Stage1 Conditional F (robust)", get_s1_condfstat_vec(res_list, robust=TRUE))))
    } else{
      extra_lines <- c(extra_lines,
                       # list(c("Stage1 F", get_s1_fstat_vec(res_list, robust=FALSE))),
                       list(c("Stage1 F (robust)", get_s1_fstat_vec(res_list, robust=TRUE))),
                       # list(c("Stage1 Conditional F", get_s1_condfstat_vec(res_list, robust=FALSE))),
                       list(c("Stage1 Conditional F (robust)", get_s1_condfstat_vec(res_list, robust=TRUE))))
    }
  # Generate latex table 
    stargazer(
      # Ordered list of results 
        res_list,
      # Add robust SE 
        se=map(res_list, calc_single_se),
      # Show the number of observations
        nobs=TRUE,
      # Add X+D labels 
        covariate.labels = c("Disaster Length","Deaths","Injuries","News Coverage"),
      # Standardize column labels   
        column.labels = as.character(c(1:5)),
      # Don't seperate columns(?) 
        column.separate = NULL,
      # Label Y
        dep.var.labels = rep("Aid",5),
      # Add lines with 1st Stage, FE and Cluster Options 
        add.lines=extra_lines,
      # Specify output type (latex for final table)
        type=output_type,
      # Omit statistics 
        omit.stat=c("ser", "adj.rsq", "rsq"),
      # Add notes
        notes.append = TRUE,
        notes=c("The Stage1 F test and Stage1 Conditional F test are weak","instrument tests reported by the felm() function in R.")
    )
}

# p <- list("dat"    = dat,
#           "y"        = "match_fema_b0",
#           "z"        = c("np_event_lead20"),
#           "d"        = "nyt_hit_flag_lead20_st",
#           "x_vars"   = c("days_of_disaster","deaths_total","injuries_total"),
#           "geo_fe"   = c("STATE_FIPS_2", "FIPS"),
#           "time_fe"  = c("year", "month"),
#           "cluster_var" = c("Date", "STATE_FIPS_2"))
# 
# gen_general_table(p)
# 
# 
# p <- list("dat"    = dat,
#           "y"        = "match_fema_b0",
#           "z"        = c("np_olympics_lead20","np_sport_lead20","np_death_lead20"),
#           "d"        = "nyt_hit_flag_lead20_st",
#           "x_vars"   = c("days_of_disaster","deaths_total","injuries_total"),
#           "geo_fe"   = c("STATE_FIPS_2", "FIPS"),
#           "time_fe"  = c("year", "month"),
#           "cluster_var" = c("Date", "STATE_FIPS_2"))
# gen_general_table(p)
```

### 2.1.6. Gen S1 Table (stage 1)
```{r}
gen_general_stage1_table <- function(p, output_type="text", z_labels=c("News Pressure"))
{
  # Generate IV results
    res_list <- gen_general_table_contents(p)
  # Extract stage1 regression results  
    res_list <- lapply(res_list, function(x) x$stage1)
  # Generate latex table 
    stargazer(
      # Ordered list of results 
        res_list,
      # Add robust SE 
        se=map(res_list, calc_single_se),
      # Show the number of observations
        nobs=TRUE,
      # Add X+Z labels 
        covariate.labels = c("Disaster Length","Deaths","Injuries",z_labels),
      # Standardize column labels   
        column.labels = as.character(c(1:5)),
      # Don't seperate columns(?) 
        column.separate = NULL,
      # Label Y
        dep.var.labels = rep("News Coverage",5),
      # Add lines with 1st Stage, FE and Cluster Options 
        add.lines=
          list(c("Covariates",  "", "", "X", "X", "X"),
               c("Time FE",     "", "X", "", "X", "X"), 
               c("Geo FE",      "", "X", "", "X", "X"),
               c("Cluster SE",  "", "", "", "", "X")),
      # Specify output type (latex for final table)
        type=output_type,
      # Omit statistics 
        omit.stat=c("ser")      
    )
}

# p <- list("dat"    = dat,
#           "y"        = "match_fema_b0",
#           "z"        = c("np_olympics_lead20","np_sport_lead20","np_death_lead20"),
#           "d"        = "nyt_hit_flag_lead20_st",
#           "x_vars"   = c("days_of_disaster","deaths_total","injuries_total"),
#           "geo_fe"   = c("STATE_FIPS_2", "FIPS"),
#           "time_fe"  = c("year", "month"),
#           "cluster_var" = c("Date", "STATE_FIPS_2"))
# 
# gen_general_stage1_table(p, z_labels=c("Olympics", "Sports", "Deaths"))

```

### 2.1.7. Gen Y~D Reg Results
```{r}
gen_general_table_contents_YonD <- function(p)
{
  # IV Only (1)
    lm1 <- p[["dat"]] %>% 
      run_YonD_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = NULL,
                   geo_fe  = NULL,
                   time_fe = NULL )
  # Add FE (3)
    lm3 <- dat %>% 
      run_YonD_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = NULL,
                   geo_fe  = p[["geo_fe"]],
                   time_fe = p[["time_fe"]])
  # Add X (2)
    lm2 <- dat %>%
      run_YonD_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = p[["x_vars"]],
                   geo_fe  = NULL,
                   time_fe = NULL)  
  # Add FE + X (5)
    lm5 <- dat %>%
      run_YonD_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = p[["x_vars"]],
                   geo_fe  = p[["geo_fe"]],
                   time_fe = p[["time_fe"]])
  # Add FE + X + Clustering (7)
    lm7 <- dat %>%
      run_YonD_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = p[["x_vars"]],
                   geo_fe  = p[["geo_fe"]],
                   time_fe = p[["time_fe"]],
                   cluster_var = p[["cluster_var"]]) 
    
  return(list("lm1"=lm1, "lm2"=lm3, "lm3"=lm2, "lm4"=lm5, "lm5"=lm7))
}
```

### 2.1.7. Gen Y~Z Reg Results
```{r}
gen_general_table_contents_YonZ <- function(p)
{
  # IV Only (1)
    lm1 <- p[["dat"]] %>% 
      run_YonZ_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = NULL,
                   geo_fe  = NULL,
                   time_fe = NULL )
  # Add FE (3)
    lm3 <- dat %>% 
      run_YonZ_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = NULL,
                   geo_fe  = p[["geo_fe"]],
                   time_fe = p[["time_fe"]])
  # Add X (2)
    lm2 <- dat %>%
      run_YonZ_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = p[["x_vars"]],
                   geo_fe  = NULL,
                   time_fe = NULL)  
  # Add FE + X (5)
    lm5 <- dat %>%
      run_YonZ_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = p[["x_vars"]],
                   geo_fe  = p[["geo_fe"]],
                   time_fe = p[["time_fe"]])
  # Add FE + X + Clustering (7)
    lm7 <- dat %>%
      run_YonZ_fun(y = p[["y"]], z = p[["z"]], d = p[["d"]],
                   x_vars  = p[["x_vars"]],
                   geo_fe  = p[["geo_fe"]],
                   time_fe = p[["time_fe"]],
                   cluster_var = p[["cluster_var"]]) 
    
  return(list("lm1"=lm1, "lm2"=lm3, "lm3"=lm2, "lm4"=lm5, "lm5"=lm7))
}
```

### 2.1.8. Gen Y~D Table
```{r}
gen_general_YonD_table <- function(p, output_type="text", d_labels=c("News Coverage"))
{
  # Generate IV results
    res_list <- gen_general_table_contents_YonD(p)
  # Generate latex table 
    stargazer(
      # Ordered list of results 
        res_list,
      # Add robust SE 
        se=map(res_list, calc_single_se),
      # Show the number of observations
        nobs=TRUE,
      # Add X+Z labels 
        covariate.labels = c("Disaster Length","Deaths","Injuries",d_labels),
      # Standardize column labels   
        column.labels = as.character(c(1:5)),
      # Don't seperate columns(?) 
        column.separate = NULL,
      # Label Y
        dep.var.labels = rep("Aid",5),
      # Add lines with 1st Stage, FE and Cluster Options 
        add.lines=
          list(c("Covariates",  "", "", "X", "X", "X"),
               c("Time FE",     "", "X", "", "X", "X"), 
               c("Geo FE",      "", "X", "", "X", "X"),
               c("Cluster SE",  "", "", "", "", "X")),
      # Specify output type (latex for final table)
        type=output_type,
      # Omit statistics 
        omit.stat=c("ser")      
    )
}
```

### 2.1.8. Gen Y~Z Table
```{r}
gen_general_YonZ_table <- function(p, output_type="text", z_labels=c("News Pressure"))
{
  # Generate IV results
    res_list <- gen_general_table_contents_YonZ(p)
  # Generate latex table 
    stargazer(
      # Ordered list of results 
        res_list,
      # Add robust SE 
        se=map(res_list, calc_single_se),
      # Show the number of observations
        nobs=TRUE,
      # Add X+Z labels 
        covariate.labels = c("Disaster Length","Deaths","Injuries",z_labels),
      # Standardize column labels   
        column.labels = as.character(c(1:5)),
      # Don't seperate columns(?) 
        column.separate = NULL,
      # Label Y
        dep.var.labels = rep("Aid",5),
      # Add lines with 1st Stage, FE and Cluster Options 
        add.lines=
          list(c("Covariates",  "", "", "X", "X", "X"),
               c("Time FE",     "", "X", "", "X", "X"), 
               c("Geo FE",      "", "X", "", "X", "X"),
               c("Cluster SE",  "", "", "", "", "X")),
      # Specify output type (latex for final table)
        type=output_type,
      # Omit statistics 
        omit.stat=c("ser")      
    )
}
```


### 2.1.9. Generate Inputs for Main Model (w/ lead Z and D vars)
```{r}
gen_inputs_wlead <- function(dat, lead_period, agg_lvl,
                             d_stems = "nyt_hits_total", #nyt_hits_total, nyt_hit_flag
                             fema_buffer = 0,
                             fema_sub = FALSE,
                             z_stems = "np_event",                       
                             subset_by_severity = FALSE,
                             nyt_agg="st",
                             binary_z_override = NULL, #by default is binary when D is binary
                             cont_zd_avg = FALSE #make D and Z avg over lead period
                             )
{
  # Generate variable names
    # FE
      if(agg_lvl=="st") geo_fe <- c("STATE_FIPS_2")
      if(agg_lvl=="cz") geo_fe <- c("STATE_FIPS_2", "FIPS")
      time_fe <- c("year", "month") 
    # Cluster
      cluster_var <- c("Date", "STATE_FIPS_2")
    # X
      x_vars <- c("days_of_disaster","deaths_total","injuries_total")      
    # Y
      if(fema_sub==FALSE) y <- paste0("match_fema_b",fema_buffer)
      if(fema_sub==TRUE)  y <- paste0("match_fema_d_b",fema_buffer)
    # Z
      if(cont_zd_avg==TRUE) dat %<>% mutate_at(vars(contains("np_event")), funs(./lead_period))
      z <- paste0(z_stems,"_lead",lead_period)
      if(d_stems=="nyt_hit_flag") binary_z <- TRUE
      if(d_stems!="nyt_hit_flag") binary_z <- FALSE
      if(!is.null(binary_z_override)) binary_z <- binary_z_override
      if(binary_z==TRUE) dat %<>% mutate_at(vars(contains("np_event")), funs(ifelse(.>0,1,0)))
    # D
      if(cont_zd_avg==TRUE) dat %<>% mutate_at(vars(contains("nyt_hit")), funs(./lead_period))
      d <- paste0(d_stems,"_lead",lead_period,"_",nyt_agg)
  # Subset by severity?
    if(subset_by_severity==TRUE){
      dat %<>%
        filter(deaths_total < 5) %>% #FOLLOW UP: make sure these cutoffs make sense
        filter(injuries_total < 100)
    }    
  return(list("dat" = dat,
              "y" = y,
              "d" = d,
              "z" = z,
              "x_vars" = x_vars,
              "geo_fe" = geo_fe,
              "time_fe" = time_fe,
              "cluster_var" = cluster_var))
}
```

### 2.1.10. Generate Inputs for 1to1
```{r}
gen_inputs_1to1 <- function(dat, lead_period, agg_lvl,
                            d_stems = "nyt_hits_total", #nyt_hits_total, nyt_hit_flag
                            fema_sub = FALSE,
                            z_stems = "np_event",                       
                            subset_by_severity = FALSE,
                            nyt_agg="st",
                            binary_z_override = NULL, #by default is binary when D is binary
                            cont_zd_avg = FALSE #make D and Z avg over lead period
                            )
{
  # Generate variable names
    # FE
      if(agg_lvl=="st") geo_fe <- c("STATE_FIPS_2")
      if(agg_lvl=="cz") geo_fe <- c("STATE_FIPS_2", "FIPS")
      time_fe <- c("year", "month") 
    # Cluster
      cluster_var <- c("Date", "STATE_FIPS_2")
    # X
      x_vars <- c("days_of_disaster","deaths_total","injuries_total")      
    # Y
      if(fema_sub==TRUE) y <- "match_fema_sub"
      if(fema_sub==FALSE) y <- "match_fema"  
    # Z
      if(cont_zd_avg==TRUE) dat %<>% mutate_at(vars(contains("np_event")), funs(./lead_period))
      z <- paste0(z_stems,"_lead",lead_period)
      if(d_stems=="nyt_hit_flag") binary_z <- TRUE
      if(d_stems!="nyt_hit_flag") binary_z <- FALSE
      if(!is.null(binary_z_override)) binary_z <- binary_z_override
      if(binary_z==TRUE) dat %<>% mutate_at(vars(contains("np_event")), funs(ifelse(.>0,1,0)))
    # D
      if(cont_zd_avg==TRUE) dat %<>% mutate_at(vars(contains("nyt_hit")), funs(./lead_period))
      d <- paste0(d_stems,"_lead",lead_period,"_",nyt_agg)
  # Subset dat based on lead period 
    if(fema_sub==TRUE){
      dat %<>%
        # Create var with days before declaration 
          mutate(days_b4_decl = as.numeric(date_declaration_sub - new_date_beg)) %>%
        # Remove rows with declaration dates before the end of the selected lead period
          filter((match_fema_sub==1 & days_b4_decl > lead_period) | (match_fema_sub==0))
    } else{
      dat %<>%
        # Create var with days before declaration 
          mutate(days_b4_decl = as.numeric(date_declaration - new_date_beg)) %>%
        # Remove rows with declaration dates before the end of the selected lead period
          filter((match_fema==1 & days_b4_decl > lead_period) | (match_fema==0))
    }
  # Subset by severity?
    if(subset_by_severity==TRUE){
      dat %<>%
        filter(deaths_total < 5) %>% #FOLLOW UP: make sure these cutoffs make sense
        filter(injuries_total < 100)
    }    
  return(list("dat" = dat,
              "y" = y,
              "d" = d,
              "z" = z,
              "x_vars" = x_vars,
              "geo_fe" = geo_fe,
              "time_fe" = time_fe,
              "cluster_var" = cluster_var))
}
```


### 2.1.11. Gen Table for Main Model (w/ lead Z and D vars)
```{r}
gen_table_wlead <- function(dat, lead_period, agg_lvl,
                            d_stems = "nyt_hits_total", #nyt_hits_total, nyt_hit_flag
                            fema_buffer = 0,
                            fema_sub = FALSE,
                            z_stems = "np_event",                       
                            subset_by_severity = FALSE,
                            nyt_agg="st",
                            binary_z_override = NULL,
                            cont_zd_avg = FALSE, #make D and Z avg over lead period
                            output_type="text")
{
  # Generate inputs
    p <- dat %>% gen_inputs_wlead(lead_period=lead_period, agg_lvl=agg_lvl,
                                  d_stems=d_stems, fema_buffer=fema_buffer,
                                  fema_sub=fema_sub, z_stems=z_stems,
                                  subset_by_severity=subset_by_severity,
                                  nyt_agg=nyt_agg,binary_z_override=binary_z_override,
                                  cont_zd_avg=cont_zd_avg)
  # Generate table   
    p %>% gen_general_table(output_type=output_type)
}

```

### 2.1.12. Gen Table for 1to1
```{r}
gen_table_1to1 <- function(dat, lead_period, agg_lvl,
                           d_stems = "nyt_hits_total", #nyt_hits_total, nyt_hit_flag
                           fema_sub = FALSE,
                           z_stems = "np_event",                       
                           subset_by_severity = FALSE,
                           nyt_agg="st",
                           binary_z_override = NULL,
                           cont_zd_avg = FALSE, #make D and Z avg over lead period
                           output_type="text")
{
  # Generate inputs
    p <- dat %>% gen_inputs_1to1(lead_period=lead_period, agg_lvl=agg_lvl,
                                  d_stems=d_stems, 
                                  fema_sub=fema_sub, z_stems=z_stems,
                                  subset_by_severity=subset_by_severity,
                                  nyt_agg=nyt_agg,binary_z_override=binary_z_override,
                                 cont_zd_avg=cont_zd_avg)
  # Generate table   
    p %>% gen_general_table(output_type=output_type)
}

# agg_lvl = "cz" 
# fema_sub <- TRUE
# dat <- load_dat_for_analysis_1to1(agg_lvl=agg_lvl, fema_sub=fema_sub)
# 
# gen_table_1to1(dat, agg_lvl=agg_lvl,
#                 lead_period=10, 
#                 fema_sub = fema_sub,
#                 d_stems = "nyt_hits_total",
#                 z_stems = "np_event")
# 
# gen_inputs_1to1(dat, agg_lvl=agg_lvl,
#                 lead_period=10, 
#                 fema_sub = fema_sub,
#                 d_stems = "nyt_hits_total",
#                 z_stems = "np_event") %>%
#   gen_general_YonD_table()

```


## 2.2. Compare: Main Model (w/ lead Z and D vars)
```{r}
agg_lvl = "cz" 
dat <- load_dat_for_analysis(agg_lvl=agg_lvl)

### Base Model
gen_inputs_wlead(dat, agg_lvl=agg_lvl,
                lead_period=10, 
                fema_sub = TRUE,
                d_stems = "nyt_hit_flag",
                z_stems = "np_event",
                cont_zd_avg=FALSE) %>%
  gen_general_YonD_table(output_type="latex")

gen_table_wlead(dat, agg_lvl=agg_lvl,
                lead_period=10, 
                fema_sub = TRUE,
                d_stems = "nyt_hit_flag",
                z_stems = "np_event",
                output_type="latex",
                cont_zd_avg=FALSE)
```

```{r}
### Break apart instrument
p1 <- gen_inputs_wlead(dat, agg_lvl=agg_lvl,
                lead_period=20, 
                fema_sub = TRUE,
                d_stems = "nyt_hit_flag",
                z_stems = c("np_olympics","np_sport_w","np_sport_s","np_death"),
                cont_zd_avg=FALSE) 

z_lab <- c("Olympics", "World Series", "Super Bowl", "Deaths")

p1 %>% gen_general_table(output_type="latex")

p1 %>% gen_general_stage1_table(output_type="latex", z_labels=z_lab)
 
p1 %>% gen_general_YonZ_table(output_type="latex", z_labels=z_lab)

```


```{r}
### Continuous: Total
gen_inputs_wlead(dat, agg_lvl=agg_lvl,
                lead_period=10, 
                fema_sub = TRUE,
                d_stems = "nyt_hits_total",
                z_stems = "np_event",
                cont_zd_avg=FALSE) %>%
  gen_general_YonD_table(output_type="latex")

gen_table_wlead(dat, agg_lvl=agg_lvl,
                lead_period=10, 
                fema_sub = TRUE,
                d_stems = "nyt_hits_total",
                z_stems = "np_event",
                output_type="latex",
                cont_zd_avg=FALSE)
```

```{r}
### Continuous: Average
gen_inputs_wlead(dat, agg_lvl=agg_lvl,
                lead_period=10, 
                fema_sub = TRUE,
                d_stems = "nyt_hits_total",
                z_stems = "np_event",
                cont_zd_avg=FALSE) %>%
  gen_general_YonD_table(output_type="latex")

gen_table_wlead(dat, agg_lvl=agg_lvl,
                lead_period=10, 
                fema_sub = TRUE,
                d_stems = "nyt_hits_total",
                z_stems = "np_event",
                output_type="latex",
                cont_zd_avg=FALSE)
```

```{r}
### Lead=20
gen_inputs_wlead(dat, agg_lvl=agg_lvl,
                lead_period=20, 
                fema_sub = TRUE,
                d_stems = "nyt_hit_flag",
                z_stems = "np_event",
                cont_zd_avg=FALSE) %>%
  gen_general_YonD_table(output_type="latex")

gen_table_wlead(dat, agg_lvl=agg_lvl,
                lead_period=20, 
                fema_sub = TRUE,
                d_stems = "nyt_hit_flag",
                z_stems = "np_event",
                output_type="latex",
                cont_zd_avg=FALSE)
```

```{r}
### Lead=40
gen_inputs_wlead(dat, agg_lvl=agg_lvl,
                lead_period=40, 
                fema_sub = TRUE,
                d_stems = "nyt_hit_flag",
                z_stems = "np_event",
                cont_zd_avg=FALSE) %>%
  gen_general_YonD_table(output_type="latex")

gen_table_wlead(dat, agg_lvl=agg_lvl,
                lead_period=40, 
                fema_sub = TRUE,
                d_stems = "nyt_hit_flag",
                z_stems = "np_event",
                output_type="latex",
                cont_zd_avg=FALSE)
```

```{r}
### Lead=60
gen_inputs_wlead(dat, agg_lvl=agg_lvl,
                lead_period=60, 
                fema_sub = TRUE,
                d_stems = "nyt_hit_flag",
                z_stems = "np_event",
                cont_zd_avg=FALSE) %>%
  gen_general_YonD_table(output_type="latex")

gen_table_wlead(dat, agg_lvl=agg_lvl,
                lead_period=60, 
                fema_sub = TRUE,
                d_stems = "nyt_hit_flag",
                z_stems = "np_event",
                output_type="latex",
                cont_zd_avg=FALSE)
```
