---
title: 'Data Analysis in R: Lecture Notes'
author: "PS239T"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    df_print: paged
    theme: flatly
    highlight: pygments
---


# 0) NOTES

## 0.1. Purpose of document 

Work in small groups to practice the data cleaning skills that we went over in class today. 

Use this opportunity to practice the skills that you will need to complete the data cleaning homework. 

If your group wishes to use a different dataset and get a head start on the homework, feel free to use a different dataset. Otherwise, I encourage you to practice by adding variables to the Florida police shootings dataset as instructed below. 


## 0.2. Instructions for data cleaning homework 

*General formatting*

Summary statistics should be accompanied by a descriptive column name and a brief explanation of the output.

Tables should be carefully formatted and polished. Use kable() and other functions to make them look ready for a conference presentation or journal. 

Comment! Remember to include at least headings and topic sentences as well as any details that you might have trouble remembering later. See Ed Rubin's website (Links to an external site.) for examples of beautiful commenting. 

*1. Find data*

Data should be at least 100 rows. 

If you download the data directly into R, make sure to save a copy of the raw data. If the data source changes, you want to always keep a copy of the raw data.

*2. Look at data using glimpse() and similar functions.*

Make sure to print a list of the variable names and data types. 

*3. Create five or more variables using mutate().*

Make sure that these variables include at least one of each data type (character, logical, numeric, and integer) as well as at least one factor variable. 

For at least one variable, show how you would create the variable using base r instead of dplyr. 

Prove that you created each of the required data types using class() or glimpse()

*4. Check missing data. Make sure that your missing data is coded correctly.* 

For example, verify that empty cells in character vectors ("") are coded as missing. 

*5. Generate a summary statistic for at least five variables using summarize().*

Make sure that these variables include at least one of each data type (character, logical, numeric, and integer) as well as at least one factor variable. 

For at least one variable, show how you would generate the same results using base r instead of dplyr. 

For at least one variable, use group_by() to calculate the statistic for different groups in your dataset. You may want to use spread() to make the output look better in a table. 

*6. Reshape your data using either gather() or spread(). Then use the other one to reshape your data back into its original  structure.*

Use head() before reshaping and after each reshaping function. We need to see what the data looks like before and after each function is used. 

*7. Save final version of cleaned data that is ready for plotting next week.* 

*8. (Optional Challenge)* Merge two or more datasets together using one of dplyr's join functions or the data.table package.

*9. (Optional Challenge)* Create a function to combine one or more of the above steps. For example, you could write a function that calculates the mean of a variable by year, uses spread() to make the years across the top of the table, and uses kable() to make the table look good.

*Hint:* If you want to include a variable name in the list of function inputs, you'll have to input the variable name as a string (as far as I know). To make dplyr recognize a variable name from a string, you can either use special dplyr functions (look up what an underscore after the function name means) or use the eval(parse(text="add command here")) trick from class. 



**********************************************************

# 1) SET UP 

## 1.1. General 

```{r}
# remove all objects
rm(list=ls())

# Unload all packages 
library(pacman)
pacman::p_unload(all)

# Add packages 
pacman::p_load(
  tidyverse, #dplyr, readr, etc.
  data.table, #fread() 
  foreign, #load data types including stata .dta files 
  magrittr, #%<>% operator
  skimr, #for summerising
  lubridate, #dates
  kable, #pretty tables
  stargazer, #pretty tables 
  xtable #pretty tables 
)
```

## 1.2. Load Data 

```{r}
load("../data/fl_shooting_data_clean.RData")
```

**********************************************************

2) ADD DATA

Download and add additional data from the following datasets and add to the cleaned data: 

* cops.csv: List of every police officer that shot someone
* tagsbycases.csv: List of every circumstance associated with an individual case
* tagsdescriptions.csv: List of circumstances by id

## 2.1. Download Data 
```{r}

```

## 2.2. Save Raw Data
```{r}

```

## 2.3. Standardize data

### 2.3.1. Look at data 
```{r}

```

### 2.3.2. Standardize vars for merging
```{r}

```

## 2.4. Combine datasets with cleaned dataset from file 02
```{r}

```

## 2.5. Save new combined file 
```{r}

```


**********************************************************

3) CLEAN NEW DATA

## 3.1. Clean the data and add variables for summarising 
```{r}

```

**Challenge:** Try to use one of the "_all", "_at", and "_if" dplyr functions. 

## 3.2. Save cleaned dataset 
```{r}

```


**********************************************************

4) SUMMARISE DATA

**Challenge:** Write a function for generating summary tables 

```{r}

```



